EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-04-02 edipack2.tex
Updated draft.

\usepackage{stackengine}
\newlength\llength
\llength=1.38ex\relax
\usepackage{xcolor}

@@

\title{EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems}
\author[a,b,c]{L.~Crippa}
\author[a]{I.~Krivenko}
\author[d]{S.~Giuli}
\author[d]{G.~Bellomia}
\author[e]{A.~Scazzola}
\author[f]{F.~Petocchi}
\author[g]{G.~Mazza}
\author[h]{L.~de Medici}
\author[d]{M.~Capone}
\author[i]{A.~Amaricci}

@@

\address[a]{\ITPHamburg}
\address[b]{\CTQMAT}
\address[c]{\WZBURG}
\address[d]{\SISSA}
\address[e]{Politecnico di Torino, Turin, Italy}
\address[f]{\Geneve}
\address[g]{\UPISA}
\address[h]{\ESPCI}
\address[i]{\CNRIOM}
% \cortext[AA]{Corresponding author:lorenzo.crippa@uni-hamburg.de}
\cortext[BB] {Corresponding author:giuli@sissa.it}
\cortext[CC]{Corresponding author:bellomia@sissa.it}

@@

Finally, the electron-phonon part of the quantum impurity problems is
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_{ph}&=\sum_q \omega_{0q} b_q^+b_q\\
    % 
    \hat{H}_{e-ph} &= \sum_q\sum_{\a\b\sigma} g_{\a\b} d^+_{\a\sigma}d_{\b\sigma}(b_q+b_q^+)
\end{split}
\end{equation}
where $q=1,\dots,N_q$ indexes the number of local phonons, $b_q$
($b_q^+$) are the destruction (creation) operators for the phonon $q$
with frequency $\omega_{0q}$. The matrix  $g_{\a\b}$ expresses is the electron-phonon coupling. 
Although feasible, dealing with more than one phonon mode becomes
quickly computationally very demanding, thus in the rest of the this
work we shall consider $N_q=1$.

In the following we consider a bath discretized into a  number
of bath degrees of freedom and a finite number of available phonons,
to cut-off the unbounded dimensions of the local phonons Hilbert space.

@@

\subsection{Classes}
The use of suitable objects enormously simplifies the  implementation
of crucial mathematical concepts aking to the diagonalization of the
quantum impurity problems. Here we discuss three main classes which
are used in the code.    
{\color{red} Text to be revised}

@@

\subsubsection{Eigenspace}\label{CodeEigenspace}
This class, contained in \texttt{ED\_EIGENSPACE} implements a
dedicated storage for the eigenvalues and eigenvectors of the quantum impurity Hamiltonian
The class defines the object \texttt{sparse\_espace}, an  
ordered linked list storing the eigenvalue (the sorting key), the
eigenvector and the corresponding QNs. To save memory the eigenvectors
are automatically distributed to all processors in shares of the right
size according to the nature of the quantum numbers $\vec{Q}$.

@@

\paragraph{{\tt bath\_type}={\bf replica}/{\bf general}} A more flexible
parametrization of the bath is represented by this topology. The
original idea of this approach is to give to each bath element a structure which {\it replicates}
that of the impurity while keeping a diagonal coupling between bath
elements and impurity. This offloads the difficulties related to the
representation and optimization of strucured quantum impurities to the
bath Hamiltonan rather than the hybridizations.
Taking a sligthly more general point of view we consider a user
defined matrix basis $\{ \Gamma^\nu \}_{\nu=1,\dots,N_{sym}}$ in the (Nambu-)spin-orbital
space and parametrize any bath Hamiltonian as:
$$
h^p_{\a\b\s\s'} = \sum_{\nu=1}^{N_{sym}} \lambda^p_\nu \Gamma^\nu_{\a\b\s\s'}
$$
where $\lambda^p_\nu\in\RRR$ are suitable variational parameters. 
The choice of the matrix basis can be inspired either by the internal structure
of the quantum impurity, i.e. $\hat{h}^0$, or be determined
case-by-case by the properties of the problem.
The total number of electronic levels used to describe the bath is in
this case: $N_b = N_{sym} N_{bath}$. 
For the {\bf replica}  topology the coupling between the impurity and
each bath elements is diagonal in the spin, orbital and internal bath
structure: $V^p_{\a\b\s\s'} = V^p$. The
{\bf general} setup introduces a generalization introducing a
dependence on the internal spin and orbital indices: $V^p_{\a\b\s\s'}
= V^p_{\a\s}\delta_{\a\b}\delta_{\s\s'}$.

All the procedures concerning the bath are grouped into a set of
modules wrapped by {\tt ED\_BATH}.

* get dimension

* replica/general bath setup

\subsection{Lanczos based Diagonalization}\label{sSecHam}

\subsection{Dynamical correlation functions}\label{sSecGF}

\subsection{Observables}\label{sSecObc}

\subsection{Reduced impurity density matrix}\label{sSecRDM}

>>>>

\usepackage{stackengine}
\newlength\llength
\llength=1.38ex\relax
\usepackage{xcolor}
\usepackage{makecell}

@@

\title{EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems}
\author[a]{L.~Crippa}
\author[a]{I.~Krivenko}
\author[b]{S.~Giuli}
\author[b]{G.~Bellomia}
% \author[c]{A.~Scazzola}
\author[d]{F.~Petocchi}
\author[e]{G.~Mazza}
% \author[f]{L.~de Medici}
\author[b]{M.~Capone}
\author[g]{A.~Amaricci}

@@

\address[a]{\ITPHamburg}
% \address[b]{\CTQMAT}
% \address[c]{\WZBURG}
\address[b]{\SISSA}
\address[c]{Politecnico di Torino, Turin, Italy}
\address[d]{\Geneve}
\address[e]{\UPISA}
\address[f]{\ESPCI}
\address[g]{\CNRIOM}
% \cortext[AA]{Corresponding author:lorenzo.crippa@uni-hamburg.de}
% \cortext[BB] {Corresponding author:giuli@sissa.it}
% \cortext[CC]{Corresponding author:bellomia@sissa.it}

@@

Finally, the electron-phonon part of the quantum impurity problems is
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_{ph}&=\sum_q \omega_{0q} b_q^+b_q\\
    % 
    \hat{H}_{e-ph} &= \sum_q\sum_{\a\b\sigma} g_{\a\b} d^+_{\a\sigma}d_{\b\sigma}(b_q+b_q^+)
\end{split}
\end{equation}
where $q=1,\dots,N_q$ indexes the number of local phonons, $b_q$
($b_q^+$) are the destruction (creation) operators for the phonon $q$
with frequency $\omega_{0q}$. The matrix  $g_{\a\b}$ expresses is the electron-phonon coupling. 
Although feasible, dealing with more than one phonon mode becomes
quickly computationally very demanding, thus in the rest of the this
work we shall consider $N_q=1$. 
In the following we consider a bath discretized into a  number
of bath degrees of freedom and a finite number of available phonons $N_{ph}$,
to cut-off the unbounded dimensions of the local phonons Hilbert space.

\paragraph{{\bf Remark}} As extensively discussed in
\onlinecite{amaricci2022}, the inclusion of local phonons with
truncated dimensions ultimately amounts to a sequential application of
the procedures defined for the electronic part, i.e. one per phonon
mode. This includes a linear scaling of the dimensions with $N_{ph}$
and of course largely limit the available degrees of freedom to
describe electronic states. In the rest of the paper we focus
specifically on the electronic part of the quantum impurity problem,
referring to the phonons in specific sections when their presence
introduces non-trivial modifications.

@@

\subsection{Classes}
The use of suitable objects enormously simplifies the  implementation
of crucial mathematical concepts aking to the diagonalization of the
quantum impurity problems. Here we discuss three main classes which
are used in the code.

@@

\subsubsection{Sparse map}\label{CodeSparseMap}
As it will be evident in the following \secu{sSecRDM}, there are cases in
which the sector construction requires to store, for any sector state,
disjoint information about the corresponding Fock state.
Let's consider a Fock state $\ket{I}$. Its bit decomposition
$\ket{\vec{n}}$ can be further split into chunks of bit corresponding,
for instance to impurity and bath degrees of freedom as 
$\ket{\vec{n}}=\ket{\vec{i}_\up\,\vec{b}_\up\,\vec{i}_\dw\,\vec{b}_\dw}$.  
The module {\tt ED\_SPARSE\_MAP} implements an hash-table {\tt
  sparse\_map}, which is included as an element in the {\tt sector} object and  
which stores  for every impurity configurations $\vec{i}_\s$ (key) the
bath states $\vec{b}_\s$ (values) corresponding to it.
The {\tt sparse\_map} ${\cal S}$ elements are constructed upon call in {\tt
  build\_sector} for any value of {\tt ed\_mode} using different
algorithms.
In the {\bf normal} case, the sector object contains two {\tt
  sparse\_maps} ${\cal S}_{\s=\up,\dw}$, in line with the
spin dependent factorization of the Fock and sector states: 
$\ket{J}=\ket{J_\up}\otimes\ket{J_\dw} \xleftarrow{\MM}
\ket{j_\up}\otimes\ket{j_\dw}=\ket{j}$.
The sparse maps are build as follows. 
For any spin state $\ket{J_\s}=\ket{{vec{i}_{\s}\vec{b}_{\s}}$ of the sector
the key is determined by the integer $I_\s$ corresponding to the
impurity bit set $\vec{i}_\s$. The values are given by any 
integer $B_\s$ corresponding to any bath bit set
$\vec{b_\s}$ associated with $\vec{i}_\s$.
Thus, any given combination of key-value
reconstructs an integer $J_\s$ representing the Fock states in the
given spin sector according to the rule $J_s = I_\s +
2^{N_{imp}}B_\s$, where $N_{imp}$ is the number of impurity bits.

In the {\bf superc/nonsu2} case the construction of the unique sparse
map is similar. For any Fock state $\ket{J}$ corresponding to sector
state $\ket{j}$ we consider the four integers $I_\up, B_\up, I_\dw,
B_\dw$ corresponding, respectively, to the bit sets
$\vec{i}_\up,\vec{b}_\up,\vec{i}_\dw,\vec{b}_\dw}$ decomposition the
state. These integers fulfill the relation
$
J =I_\up +  B_\up2^{N_{imp}} + (I_\dw +  B_\dw2^{N_{imp}})2^{N_s}
$
To obtain a contiguous memory pattern we define as a key of the sparse
map the combination $I_\up + I_\dw2^{N_{imp}}$ and as corresponding
  values the integer $B_\up + B_\dw2^{N_b}$. Where needed the correct
  bit sets and Fock state can be reconstructed with simple algebra at
  no additional cost.

\setbox0=\hbox{%
  \begin{minipage}{0.5\linewidth}
    \center{\bf normal}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
i=0
do Iup=0,2**Nbit-1     
  nup_ = popcnt(Iup) 
  if(nup_ /= Nups(1))cycle
  i=i+1 
  !...
  iImp  = ibits(iup,0,Norb)
  iBath = ibits(iup,Norb,Norb*Nbath)
  call sp_insert_state(self%H(1)%sp,iImp,iBath,i)
enddo
i=0
do Idw=0,2**Nbit-1
  ndw_= popcnt(Idw)
  if(ndw_ /= Ndws(1))cycle
  i=i+1 
  !...
  iIMP  = ibits(idw,0,Norb)
  iBATH = ibits(idw,Norb,Norb*Nbath)
  call sp_insert_state(self%H(2)%sp,iImp,iBath,dim) 
enddo
\end{lstlisting}
\end{minipage}
}
\savestack{\listingA}{\box0}

\setbox0=\hbox{%
  \begin{minipage}{0.5\linewidth}
    \center{\bf superc/nonsu2}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
i=0
do Idw=0,2**Ns-1
  ndw_= popcnt(idw)
  do Iup=0,2**Ns-1
    nup_ = popcnt(iup)
    !superc 
    sz_  = nup_ - ndw_
    if(sz_ /= self%Sz)cycle
    !nonsu2 
    !nt_  = nup_ + ndw_
    !if(nt_ /= self%Ntot)cycle 
    i=i+1 
    !...
    iImpUp  = ibits(iup,0,Norb)
    iImpDw  = ibits(idw,0,Norb)
    iBathUp = ibits(iup,Norb,Norb*Nbath)
    iBathDw = ibits(idw,Norb,Norb*Nbath)
    iImp    = iImpUp  + iImpDw*(2**Norb)
    iBath   = iBathUp + iBathDw*(2**(Norb*Nbath))
    call sp_insert_state(self%H(1)%sp,iImp,iBath,dim)
  enddo
enddo

\begin{tabular}{c|c}\label{list2}
  \stackinset{l}{}{t}{}{}{\listingA} &
\stackinset{l}{}{t}{}{}{\listingB} 
\end{tabular}
%

\subsubsection{Eigenspace}\label{CodeEigenspace}
The module \texttt{ED\_EIGENSPACE} implements a list
storing for the eigenvalues and eigenvectors of the quantum impurity
Hamiltonian. 
This class defines the object \texttt{sparse\_espace}, an  
ordered linked list storing the eigenvalue (the sorting key), the
eigenvector and the corresponding QNs. To save memory the eigenvectors
are automatically distributed to all processors in shares of the right
size according to the nature of the quantum numbers $\vec{Q}$.

@@

\paragraph{{\tt bath\_type}={\bf replica}/{\bf general}} A more flexible
parametrization of the bath is represented by this topology. The
original idea of this approach is to give to each bath element a structure which {\it replicates}
that of the impurity while keeping a diagonal coupling between bath
elements and impurity. This offloads the difficulties related to the
representation and optimization of strucured quantum impurities to the
bath Hamiltonan rather than the hybridizations.
Taking a sligthly more general point of view we consider a user
defined matrix basis $\vec{\Gamma}=\{ \Gamma^\nu_{\a\b\s\s'} \}_{\nu=1,\dots,N_{sym}}$ in the (Nambu-)spin-orbital
space and parametrize any bath Hamiltonian as:
$$
h^p = \sum_{\nu=1}^{N_{sym}} \lambda^p_\nu
\Gamma^\nu\equiv \vec{\lambda}^p\cdot \vec{\Gamma}
$$
where $\vec{\lambda}^p\in\RRR^{N_{sym}}$ is a vector of variational parameters. 
The choice of the matrix basis can be inspired either by the internal structure
of the quantum impurity, i.e. $\hat{h}^0$, or be determined
case-by-case by the properties of the problem.
The total number of electronic levels used to describe the bath is in
this case: $N_b = N_{sym} N_{bath}$. 
For the {\bf replica}  topology the coupling between the impurity and
each bath elements is diagonal in the spin, orbital and internal bath
structure: $V^p_{\a\b\s\s'} = V^p$. The
{\bf general} setup introduces a generalization introducing a
dependence on the internal spin and orbital indices: $V^p_{\a\b\s\s'}
= V^p_{\a\s}\delta_{\a\b}\delta_{\s\s'}$.

All the procedures concerning the bath, either on the user side or the
internal {\tt effective\_bath}, are grouped into a set of modules
wrapped by {\tt ED\_BATH}.
We divided the set of modules in three categories, according to their
scope.
\paragraph{Bath Auxiliary Tools}
This part contains some functions directed to the user as, for
istance, the implementation of suitable conventional symmetry
operations acting on the user bath array like, e.g., orbital symmetry,
particle-hole symmetry, etc. found in {\tt ED\_BATH\_USER}
However, the key functions of this group concerns the determination
of the total dimension of the user bath array in {\tt ED\_BATH\_DIM}.
Starting from the values
of some input variables the function {\tt ed\_get\_bath\_dimension}
returns the dimension $B$ to which the  user should allocate the bath
array to contains all and just the bath parameters. Any following call
to functions hosting the bath as an input will check that the user supplied
bath array has the correct dimensions compared to the value of $B$
using {\tt check\_bath\_dimension}.

\paragraph{Bath Replica (General)}
The module {\tt ED\_BATH\_REPLICA} implements the class {\tt Hreplica}
for the  the {\bf replica} ({\bf general}) bath parametrization. Beside the
conventional operations of object construction, destructin, reading
and saving the module includes the matrix basis $\{ \Gamma^\nu
\}_{\nu=1,\dots,N_{sym}}$ and variatinal parameters 
$\vec{\lambda}^p$ setup, i.e. {\tt set\_Hreplica} ({\tt
  set\_Hgeneral}) as well as a dedicated function to build  the bath Hamiltonian $h^p =
\vec{\lambda}^p\cdot \Gamma$ in {\tt build\_Hreplica} ({\tt
  build\_Hgeneral}).

\paragraph{Effective Bath}
The module {\tt ED\_BATH\_DMFT} implements a class for the object {\tt
  effective\_bath} (already defined in {\tt ED\_VARS\_GLOBAL}). This
data structure contains and handles all the actual parameters of the
bath, for any choice of {\tt ed\_mode} and {\tt bath\_type},
referencing the class {\tt Hreplica} only for the {\bf
  replica/general} bath topology.
A global instance of the class {\tt dmft\_bath}  containing the
current bath parameters is shared throughout the code.   
The class constructor is supplemented with a specific procedure initializing
the effective bath {\tt init\_dmft\_bath} which guess the bath
parameters or read them from a file specified by the input variable {\tt
  Hfile}.
The class also contains procedures to transform the instance
{\tt dmft\_bath} to/from the user bath array: {\tt
  get/set\_dmft\_bath}.

% \begin{table}%[ht]
%   \label{TabBath}
% \begin{center}
% \begin{tabularx}{\linewidth}{ |X|X|X|X| } 
%   \hline
%   c  & {\bf normal} & {\bf superc} & {\bf  nonsu2} \\
%   \hline
%   {\bf normal} $[N_b\!=\!N_{bath}N_\a]$ & $\vec{\e}[N_b],\vec{V}[N_b]$ 
%                     & $\vec{\e}[N_b], \vec{\Delta}[N_b],\vec{V}[N_b]$
%                                    &  $\vec{\e}[N_b], \vec{V}[N_b],\vec{W}[N_b]$ \\
%   \hline
%   {\bf hybrid} $[N_b\!=\!N_{bath}N_\a]$ & $\vec{\e}[N_{bath}],\vec{V}[N_b]$ 
%                     & $\vec{\e}[N_{bath}], \vec{\Delta}[N_{bath}],\vec{V}[N_b]$
%                                    &  $\vec{\e}[N_{bath}], \vec{V}[N_b],\vec{W}[N_b]$ \\
%   \hline
%   {\bf replica} $[N_b\!=\!N_{bath}N_{sym}]$ & \multicolumn{3}{*{\hat{h} } \cr
%   \hline
%   {\bf general} $[N_b\!=\!N_{bath}N_{sym}]$ & \cline{2-4} & b&  c  \cr
%   \hline
% \end{tabularx}
% \end{center}
% \caption{Ciao}
% \end{table}

\subsection{Lanczos based Diagonalization}\label{sSecHam}
In presence of symmetries (see \secu{sSecQNs}) the matrix representing
the Hamiltonian operator $\hat{H}$ takes a block diagonal form. Each block
contains the Hamiltonian matrix $H_{\SS_{\vec{Q}}}$ for a
given symmetry sector with quantum number $\vec{Q}$.
The analysis of the energy spectrum of the quantum impurity problems
thus reduces to the sequential solution of the sector secular problem
$$
H_{\SS_{\vec{Q}}}\ket{v}=E_{\SS_{\vec{Q}}}\ket{v}.
$$

Although the dimension of any such sector $\SS$ are much smaller than the
total Fock space dimension $D_\SS \ll D_\FF$ still the complete solution of
the related eigenvalue problem represents an unsurmountable problem
already with $N_s\simeq 8$. 
In order to outwit this issue a number of
algorithms have been developed leveraging on the sparse nature of
$H_\SS$, e.g. Krylov or similar sub-space methods \cite{Lanczos,Arnoldi,Feast}.   
In \NAME the alghoritm of choice is P-ARPACK, which represents the
state-of-the-art Lanczos-based eigensolver with distributed memory
support~\cite{P-Arpack}. This method includes re-orthogonalization
within a block Lanczos algorithm which ensures convergence of the
requred eigenpairs. Using the input variable {\tt lanc\_method} it is
however possible to select a simple, but not as efficient, parallel
Lanczos algorithm.
The core of Krylov-based diagonalization algorithms is the Matrix-Vector Product
(MVP), i.e. the application of the Hamiltonian to a given input
vector:
$$
H_\SS \ket{v} \to \ket{w}\qquad \ket{v},\ket{w}\in \SS_{\vec{Q}}
$$

This operations easily takes over more than 80\% of the computation
time, so its optimization is a crucial aspect in any diagonalization
algorithms. In \onlinecite{amaricci2022} we covered in full details
some distrubuted memory massively parallel algorithms to address this
specific issue.
Here, using the same naming convention introduced in \cite{amaricci}, 
we briefly outline the main strategy for the implementation of the MVP
according to the value of {\tt ed\_mode}.

\paragraph{{\bf normal}}
This case relies on the tensor structure of the Fock space and the
symmetry sectors. In any given sector the electronic part of the
Hamiltonian matrix reads:
\begin{equation}
H_\SS = H_d  + H_\up\otimes \mathbb{I}_\dw + \mathbb{I}_\up\otimes
H_\dw + H_{nd}\,.
\label{HssNormal}
\end{equation}
where $H_d$ is a diagonal term containing the local part of the
Hamiltonian, there including the density-density terms of the
interaction,  $H_\sigma$ describes the hopping processes of the
electrons with spin $\sigma=\up,\dw$.
The term $H_{nd}$ contains all the remaining non-diagonal
elements which do not fit in the previous two components,
e.g. the spin-exchange and pair-hopping interaction terms.
Correspondingly, the any sector state $\ket{v}$ can be expressed in a matrix
basis $\hat{v}$ with rows (columns) running over spin $\up$ ($\down$)
configurations. Leveraging on this setup it is possible the MVP using the 
{\tt MPI\_All2AllV} which aims at optimizing locality in the memory
locality. Additional terms contained in $H_{nd}$ are treated using the
{\tt MPI\_AllGatherV} algorithm which includes a slight communication
overload ultimately reducing the parallel efficiency.

\paragraph{{\bf superc/nonsu2}}
In this two cases the symmetry enforce a specific relation between $\up$ and
$\dw$ configurations ultimately suppressing the benefits arising from
Fock space factorization.
The electronic part of the Hamiltonian matrix is usually organized in this structure:
$$
H_\SS = H_{imp} + H_{int} + H_{bath} + H_{imp-bath} 
$$
where, with almost obvious meaning of the labels, $H_{imp}$ encodes
the non-interacting part of the impurity Hamiltonian determined by
$h^0_{\a\b\s\s'}$, $H_{int}$ the interaction term, $H_{bath}$ the
effective bath terms related to $h^p_{\a\b\s\s'}$ and finally
$H_{imp-bath}$ describes the coupling between impurity and bath
proportional to $V^p_{\a\b\s\s'}$. 
In an MPI setup the first three terms contains elements local in the
memory of each node. These are stored separately from the non-local
ones in any {\tt sparse\_matrix} representing $H_\SS$. 
The MVP is carried out using the {\tt MPI\_AllGatherV} algorithm
which, as discussed in \onlinecite{amaricci2022}, requires the
reconstruction of the distributed input array any node producing a
overflow in the MPI communication.

The diagonalization procedure is divided in two main steps: a global
setup and the actual diagonalization.
%
The setup step takes care of allocating the required memory,
setup the MPI environment and allocate the correct 
Matrix-Vector Product (MVP) procedure according to the symmetries of
the problem. 
This part is implemented in a set of independent modules  {\tt
  ED\_HAMILTONIAN\_?}, where {\tt ed\_mode=?}.
Specifically, any such module includes two functions.
The first, {\tt  build\_Hv\_sector\_?}, builds the required sector and
allocate the MVP function, i.e. point the abstract function {\tt
  sphtimesv\_p} to the correct procedure determined by the value of
{\tt ed\_sparse\_H=True/False}.
If {\tt True} the Hamiltonian terms in $H_\SS$
get evaluated and stored in a {\tt sparse\_matrix} instance, possibly
using automatic parallel storage, and then used to perform
the MVP. This part is implemented in the modules {\tt
  ED\_HAMILTONIAN\_?\_STORED\_HxV}
If {\tt ed\_sparse\_H} is {\tt False} the MVP is operated on-the-fly,
i.e. each element of $H_\SS$ is direcrtly applied to the given input
vector $\ket{v}$ to determine the outgoing vector $H_\SS\ket{v}$,
either in serial or parallel mode as implemented in {\tt
  ED\_HAMILTONIAN\_?\_DIRECT\_HxV}.

The function {\tt vecDim\_Hv\_sector\_?} returns the 
dimension of the vector to be used in the
MVP. In serial execution this number is just $D_\SS$ the dimension of
the sector, whereas in a parallel mode the returned value is the
dimension $d_i$ of the vector chunk per each node such that
$\sum_{i=1}^{N_{nodes}}d_i = D_\SS$. Yet, the specific value of $d_i$
depends on the MPI algorithm used for the MVP function, which in turn
is determined by the symmetry of the problem, i.e. {\tt ed\_mode}.

The sector-by-sector diagonalization for each
value of {\tt ed\_mode} is implemented in the modules {\tt
  ED\_DIAG\_?} through the main functions {\tt
  diagonalize\_impurity\_?}. The diagonalization proceeds in three main
steps: i) setup the diagonalization by selecting all or a sub-set of the
sectors to analyze; ii) sequential diagonalization cycle over all the suitable symmetry
sectors; iii) analysis of the resulting {\tt state\_list} containins
all the eigensolutions to be conserved.

\subsection{Dynamical correlation functions}\label{sSecGF}
The determination of the low energy part of the Hamiltonian spectrum
enables the evaluation of dynamical correlation functions using Krylov
sub-space algorithm.
This is a key part of the library when using \NAME as an impurity
solver within DMFT framework.

Before entering into the implementation details we briefly outline the
generic algorithm. Let us consider the dynamical correlation function:
\begin{equation}
  \label{eqGaa}
  C_\AA = \ibra \TT_\pm[ \AA(t) \AA^+ ]\iket
  \end{equation}
where $\AA(t)=e^{iHt}\AA e^{-iHt}$, $\TT_\pm$ is the time-ordering operator for fermions or
bosons and $\ibra \AA \iket=\frac{1}{Z}\Tr{ \left[ e^{-\beta
      H}\AA \right]}$,  $Z=\sum_ne^{-\beta E_n}$, is the thermodynamic
average. Using spectral decomposition, the expression \equ{eqGaa} is reduced into Kallen-Lehman form:
\begin{equation}\label{KLgf}
  \begin{split}
    C_\AA(z) 
    &=  \ibra \AA \frac{1}{z-H} \AA^+\iket
    \mp \ibra \AA^+ \frac{1}{z+H} \AA\iket \cr
    %
    & =\frac{1}{Z}\sum_n e^{-\beta E_n}\sum_m
  \frac{ |\bra{\psi_m}\AA^+ \ket{\psi_n}|^2 }  {z-(E_m-E_n)}
  \mp
  \frac{ |\bra{\psi_m}\AA \ket{\psi_n}|^2 }  {z+(E_m-E_n)}\cr
\end{split}
\end{equation}
where $z\in\CCC$, 
$\ket{\psi_n}$, $E_n$ are the eigenpairs of the
Hamiltonian $H$ and $E_0$ is the groundstate energy.
While appealing, the previous expression requires the knowledge of the
entire Hamiltonian spectrum. A key simplification is obtained looking
at the first line in \equ{KLgf},
showing that the $C_\AA$ essentially corresponds to a given element of the inverse operator $(z-H)^{-1}$, i.e. the
resolvant. The Krylov sub-space method provides a controlled
approximation for this element, leveraging over the
reduction of the (sector) Hamiltonian to a partial tri-diagonal
form and returning an arbitrary number of amplitudes to (unknown)
excited states. To illustrate the algorithm let's consider the normalized initial state: 
$$
\ket{\phi_n}=\AA^+\ket{\psi_n}/\NN_n
$$
where  $\ket{\psi_n}\in\SS$ and $\NN_n=\sqrt{\bra{\psi_n}\AA
  \AA^+\ket{\psi_n}}$.
We can construct the Krylov basis $$\mathcal{K}_{N} (\ket{\phi_n})=\{\ket{\phi_n}, H\ket{\phi_n}, \dots,
H^N\ket{\phi_n}\}\equiv \{\ket{v^n_0}, \ket{v^n_1},\dots, \ket{v^n_N}
\}$$ with $1 \ll N \ll \DD_\SS$ by iteratively applying the (sector)
Hamiltonian through MVP functions. 
Any given eigenstate $\ket{\psi_n}$ has components along the Krylov basis $\mathcal{K}_{N} (\ket{\phi_n})$ which read: $\ket{\psi_n} =
\sum_i  \ibra v^n_i|\psi_n \iket  \ket{ v^n_i} =
\sum_i a^n_i \ket{v^n_i}$.   We can evaluate the operator components
$=\bra{\psi_m}\AA^+ \ket{\psi_n}\equiv \langle\psi_m| \phi_n\rangle =
\NN_n {a^{n*}_m}$. Inserting this expression into \equ{KLgf} we
obtain the following approximation to $C_\AA$:
\begin{equation}
  \label{eqGKrylov}
  \begin{split}
    C_\AA(z)  &\simeq \frac{1}{Z}\sum_n e^{-\beta E_n}
    \sum_{m=1}^{N} \frac{\bra{\psi_n}\AA \AA^+\ket{\psi_n} |a^n_m|^2}{
      z - (E_m-E_n)} \mp \frac{\bra{\psi_n}\AA^+ \AA\ket{\psi_n}
      |a^n_m|^2}{ z + (E_m-E_n)}\cr
    %
    &= \frac{1}{Z}\sum_n\sum_{m=1}^{N} \sum_{\nu=\pm}\frac{w^{\nu}_{mn}[\AA]}{z - dE^\nu_{mn}[\AA]} 
    = \frac{1}{Z}\sum_n
    \sum_{m=1}^{N} \sum_{\nu=\pm} g_\AA(\nu, w^{\nu}_{mn},  dE^\nu_{mn}) \cr
  \end{split}
\end{equation}
where we introduced the notation $g(\nu, w(\AA)^{\nu}_{mn},  dE(\AA)^\nu_{mn})$
for the {\tt gfmatrix} object containing the weights $w(\AA)^{\nu}_{mn}$ and
poles $dE(\AA)^{\nu}_{mn}$ for the operator $\AA$, for every initial state $\ket{\psi_n}$
contributing to the low energy spectrum, for every order $m$ of the
Krylov sub-space algorithm and for channel $\nu$.

A stringent limitation of this procedure is to get applied to diagonal
dynamical correlation functions of the form \equ{eqGaa}. However, as
we will show in the following, many practical application requires to
evaluate off-diagonal functions of the form:
$C_{\AA\BB}(z) = \ibra T_{\pm}[\AA(t) \BB^+]\iket$.
A direct extensin of the method is obtained considering suitable
auxiliary operators prepared with independent linear combination of
$\AA$ and $\BB$.
A notable example is to consider the two auxiiary operators $\OO=\AA +\BB$ and
$\PP=\AA-i\BB$. Using simple algebra it is then straightforward to
obtain the desired function $C_{\AA \BB}$ from the evaluation of
$C_\OO$ and $C_\PP$:
$$
C_{\AA\BB} = \frac{1}{2}\left[C_\OO + C_\PP - (1-i)C_\AA -(1-i)C_\BB\right]
$$

In \NAME the impurity Green's functions $G_{\a\b\s\s'}=\ibra \TT_\pm[ c_{\a\s}(t)
c_{\b\s'}^+ ]\iket$,
is contained in \texttt{ED\_GREENS\_NORMAL}. This wraps more
specific methods according to the symmetry of the problem expressed by
{\tt ed\_mode}. In the {\bf normal} mode also spin,
charge, pair and excitonic susceptibilities functions. 
Computationally, the construction of the Krylov basis
$\mathcal{K}_N(\OO\ket{\psi_n})$ for any eigenstate of the low energy
spectrum is the most time consuming step. As for the diagonalization
part, here a crucial speed-up is obtained by the parallel execution of
the MVP lying at the earth of the Hamiltonian tri-diaagonalization
process. The input variable {\tt lanc\_gfniter} controls the largest order of
the Krylov basis, i.e. the maximum number of excitations in
\equ{eqGKrylov}.  
Operationally, each symmetry mode {\tt ed\_mode=?} requires a different construction of
the Green's functions implemented in {\tt ED\_GF\_?}

\paragraph{{\bf normal}}
In this case the all the orbital dependent and spin-diagonal Green's
functions $G_{\a\b\s\s}$ needs to be evaluated. The orbital diagonal case
$\a=\b$ proceeds as discussed above considering $\AA=c_{\a\s}$. This
step makes use of the {\tt apply\_op\_C/CDG} functions in {\tt
  ED\_AUX\_FUNX} on any eigenstate $\ket{\psi_n}$ in the {\tt
  state\_list}, i.e. a global instance of {\tt sparse\_espace}
containing the low energy spectrum of the problem.
The sector Hamiltonian matrix in this mode is assumed to be real
symmetric. This brings a simplification in the evaluation of the
off-diagonal components $G_{\a\b\s\s}=G_{\b\a\s\s}$.
These components are thus evaluated considering the auxiliary operator
$\OO=(c_{\a\s} + c_{\b\s})$ and using the relation
$G_{\a\b\s\s}=\tfrac{1}{2}(C_\OO - G_{\a\a\s\s} - G_{\b\b\s\s})$.

The diagonal spin, charge and pair susceptibility terms
$\chi^{S^z}_{\a\a}$, $\chi^N_{\a\a}$ and $\chi^\Delta_{\a\a}$ are
evaluated employing, respectively, the operators $\AA=\sum_{\s\s'}c^+_{\a\s}\tau^z_{\s\s'}c_{\a\s'}\equiv
S^z_\a$, $\AA=\sum_{\s\s'}c^+_{\a\s}\tau^0_{\s\s'}c_{\a\s'}\equiv
N_\a$) and $\AA=c_{\a\dw}c_{\a\up}\equiv
\Delta_\a$, where
$\tau^{a=0,x,y,z}$ are the Pauli matrices.
Similarly, yhe off-diagonal terms are obtained by considering the operator $\AA=S^z_\a
+ S^z_\b$, $\AA=N_\a + N_\b$ or $\AA=\Delta_\a + \Delta_\b$. 
The excitonic susceptibility $\chi^T_{\a\b}$ is defined with respect
to the vector operator: $T^i_{\a\b} = \sum_{\s\s'}
c^+_{\a\s}\tau^i_{\s\s'}c_{\b\s'}$, respectively, the spin-singlet
($i=0$) and the spin-triplet excitons ($i=x,y,z$) operators.
